
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}
\vspace{1.5cm}

\large

% 2 sentences - Fluid simulations are important for XYZ
Using CFD programs to simulate fluids has become an incredibly important element of many research areas and industrial applications such as weather forecasting, animation, and vehicle design.
% have become a hugely important element of many design processes.
% Fluid simulations are used for weather forecasting, animation, and designing vehicles to name a few examples.
%to predict the weather, create animation, and to determine aerodynamic properties of vehicles.
% 1 sentences - Making them go fast is important.
Complex simulations may require large \& expensive systems to complete, and even then may take multiple hours to finish.
% This is unhelpful for the simulation users, who may benefit from quicker simulations to e.g. speed up their design iterations.
Making simulations run faster on relatively cheap GPU hardware would lower the barrier to entry, and help users work effectively by reducing their iteration times.

% 1 sentences - Visualizing in-situ is important.
% 1 sentences - Visualizing quickly is important. \todomark{Avoid round trip of full simulation -> visualization, which can take a long time}
% User efficiency can also be improved with in-situ visualization, where data is visualized in parallel with the simulation instead of waiting for it to complete.
% For advanced visualization features to be included, or to combine many different visualization features, an implementation must be efficient enough to not delay the simulation work.
User efficiency may also be improved with real-time in-situ visualization, where data is visualized and displayed in parallel with the simulation.
For advanced visualization features to be included, or to combine many different visualization features, an implementation must be efficient enough to not delay the simulation work.

% 2 sentences - Goal is to introduce a high-speed tightly-coupled in-situ visualization based on ACA. Novelty comes from implementing techniques from games industry, mixing Vulkan and CUDA for an in-situ visualization.
The goal of this project is to implement a 
real-time tightly-coupled in-situ visualized fluid simulation, based on the simulation code from the 2020 CS257 coursework.
While real-time in-situ visualizations exist in games, this implementation provides novelty by focusing on accuracy over graphical fidelity.
The simulation is moved to the GPU to improve performance, and optimized for the target platform.
The visualization is implemented from scratch with Vulkan, using techniques from the games industry for efficient rendering.
% The implementation is evaluated vs. the original coursework, and impact of an advanced visualization is shown to be negligible compared to the simulation on the target hardware.
The implementation is evaluated against the original simulation and other simulation programs, and the impact of an advanced visualization is shown to be negligible compared to the simulation.

\vspace{0.5cm}

\noindent \textit{Keywords: Fluid Simulation, CUDA, Vulkan, GPU, In-Situ, Visualization}