% !TEX root =  ../FinalReport.tex

\newcommand{\must}[0]{\textbf{must}}
\newcommand{\should}[0]{\textbf{should}}
\newcommand{\shouldnt}[0]{\textbf{should not}}


\chapter{Project Requirements}
\label{sec:Requirements}
Ahead of initial development, a set of requirements were created to further specify the project goals.
These requirements evolved as more research was completed, for example, the visualization-related requirements were only determined after visualization research completed (see \cref{sec:ProjectManagement}).
This chapter shows the functional and non-functional requirements, prioritized as either \must{}-have or \should{}-have.
Complex requirements have sub-requirements, which clarify certain features that must/should be present for the top-level requirement to be met.
%  which  into sub-requirements in some cases to clarify the broad points.
% Each requirement is assigned a 
The hardware and software constraints for the program are also shown.



% These basic functional and non-functional requirements define the baseline the final result will be measured against.
% \todomark{Elaborate on requirements}

% Functional Requirements define the actions a program must/should be able to perform, such as \cref{req:GenerateState} stating ``the system must be able to generate initial simulation states''.
% The functional requirements, the tests performed to check them, and the outcomes have been listed in \cref{tab:functional_req}.

% Non-functional Requirements do not define actions, but rather define properties of those actions that must be met.
% As an example \cref{reqN:SimSpeed} states that with the same initial input, the CUDA simulation ``should run at least 2x as fast as the original coursework''.


\section{Functional Requirements}
Functional Requirements define actions a program must/should be able to perform.
\begin{reqF}
    \item \label{req:StoreState} The system \must{} store simulation state in a file or set of files.
    \item \label{req:LoadState} The system \must{} be able to load the initial state of a simulation from these file(s).
    \item \label{req:GenerateState} The system \must{} be able to generate initial simulation state files.
    \item \label{req:HeadlessSim} The system \must{} be able to simulate from an initial state for a set amount of time without visualizing.
    \begin{reqF}
        \item \label[reqFi]{req:HeadlessOutput} This mode \must{} be able to store the final state to output file(s).
    \end{reqF} 
    \item \label{req:VizSim} The system \must{} be able to simulate from an initial state for an indeterminate amount of time while visualizing.
    \begin{reqF}
        \item \label[reqFi]{req:VizPauseResume} This mode \must{} allow the user to pause and resume the simulation.
        \item \label[reqFi]{req:VizSaveState} This mode \should{} be able to save it's state to output file(s) when requested.
        \item \label[reqFi]{req:VizManip} This mode \should{} allow the user to manipulate the simulation or visualization state while simulating.
        \item \label[reqFi]{req:VizLockedFPS} This mode \should{} be able to run at a locked frame-rate.
        \item \label[reqFi]{req:VizFlatOut} This mode \should{} be able to run as fast as possible, without locking the framerate.
        \item \label[reqFi]{req:VizSomeSpeed} This mode \must{} be able to perform at least one of \cref{req:VizLockedFPS,req:VizFlatOut}.
    \end{reqF} 
    \item \label{req:GPUCapable} Both methods of simulation \must{} be capable of using the GPU for simulating.
    \item \label{req:Compare} The system \must{} be able to compare how similar two simulation states are.
    \begin{reqF}
        \item \label[reqFi]{req:CompareBinary} This comparison \should{} produce a binary SIMILAR/NOT~SIMILAR verdict using heuristics.
    \end{reqF}
    
    \item \label{req:VizLayers} The visualization \must{} consist of multiple layers which can be individually controlled.
    \begin{reqF}
        \item \label[reqFi]{req:VizLayersBackground} The visualization \must{} always display a background layer which shows the simulation obstacles in a different color to the fluid.
        \item \label[reqFi]{req:VizLayersScalar} The visualization \must{} be able to display an optional scalar quantity (e.g. pressure) using a color scale, where the value is within a user-defined range.
        \item \label[reqFi]{req:VizLayersVector} The visualization \must{} be able to display an optional vector quantity (e.g. velocity) using a vector field, where the magnitude is within a user-defined range.
        \item \label[reqFi]{req:VizLayersParticle} The visualization \must{} feature an optional particle simulation, where particles are continuously emitted and move with the velocity of the field.
    \end{reqF}
    \item \label{req:VizAutoRange} The scalar and vector quantities (\cref{req:VizLayersScalar,req:VizLayersVector}) \should{} have an auto-range function to automatically calculate the range based on the values present.
    \item \label{req:VizColors} All colors used in the visualization (e.g. particle colors, the scalar color scale) \should{} be user-controlled.
    \item \label{req:VizParticleEmission} The locations of particle emitters \should{} be user-controllable.
\end{reqF}

\pagebreak
\section{Non-Functional Requirements}
Non-functional Requirements do not define actions, but rather define properties of those actions or the program that must be met.
\begin{reqNF}
    \item \label{reqN:LargeData} The system \must{} be capable of operating on large datasets (e.g. 4096x4096 grids) without failing.
    \item \label{reqN:Resources} The system \must{} be efficient and avoid wasting any resources allotted to it.
    \item \label{reqN:SimilarOutput} The simulation \must{} produce similar results to the original simulation when equivalent initial state is used.
    \item \label{reqN:SimSpeed} The simulation \should{} run at least 2x as fast as the original simulation when equivalent initial state is used.
    \item \label{reqN:Realtime} The visualized simulation \must{} run in real-time at framerates~$\ge$~30 FPS for some outputs.
    \item \label{reqN:VizSpeed} The visualization features \shouldnt{} have a significant impact on the framerate.
    \item \label{reqN:Intuitive} The visualized simulation \should{} intuitively represent the fluid flow such that it can be understood by someone unfamiliar with fluid simulation.
    \item \label{reqN:VizParticleAdvanced} The particle simulation (\cref{req:VizLayersParticle}) \should{} demonstrate advanced behaviour to make the visualization more intuitive e.g. avoiding clumping.
    \item \label{reqN:Documented} The system \must{} be fully documented and maintainable. % TODO Wheeler referenced BCS code of conduct again??
    \item \label{reqN:UsageGuide} The system \should{} have a simple guide to common operations for new users to refer to.
    \item \label{reqN:DCSCompile} The system \should{} be fully compilable and executable from a DCS machine with minimal extra installations.
\end{reqNF}

\section{Hardware and Software Constraints}
\label{sec:Requirements_HardwareSoftware}
As this simulation uses a GPU, the developer must have one available for debugging and testing the program.
As the CUDA API is used to implement the simulation (see \cref{sec:LibrarySelection}), the program requires an NVIDIA GPU to run.
Due to COVID-19 restrictions, the only hardware available to test the device was the researcher's GTX~1080, but it should function on all devices which support CUDA 10.

The high-speed rendering requirements of the program necessitated the use of Vulkan over OpenGL.
Vulkan gives the developer finer control over scheduling and allows the hardware to take shortcuts that it may not be able to do under OpenGL.
For more on this decision see \cref{sec:LibrarySelection}.