\subsection{File Formats}\label{sec:FileFormat}
To fulfil \cref{req:StoreState} two file formats have been defined to store simulation data and parameters.

\subsubsection{Fluid Parameters}
Parameters that are characteristic of a particular fluid or simulation type are stored in a ``Fluid Parameters'' file.
This includes the Reynolds number, the timestep safety factor, and the maximum iteration count for the Poisson solver.
They are stored in a JSON format to be human-readable, are reusable for different simulation states, and can be easily edited by the end user.
An example is shown in \cref{fig:FluidParamsExample}.

\input{Ch42Design/figures/example_fluidparams}

\subsubsection{Simulation State}
Data unique to an individual state such as simulation resolution, physical size, and velocity fields are stored in a binary format reused from the original simulation.
As the data is much more sensitive to individual modifications\footnote{For example, changing a single value in the velocity field can introduce discontinuities.}, it makes more sense to store this data in a binary format where it cannot be easily modified by a user.
Additionally, the binary format is much smaller than any text-based format, which helps as the volume of data stored is much larger than that stored in the fluid parameters.

The header consists of a pair of unsigned 32-bit integers specifying the resolution of the simulation, and a pair of 32-bit floating-point numbers specifying the physical dimensions of the simulation.
From there, four sets of data for each column are stored, including the boundary padding squares:
\begin{enumerate}
    \item Horizontal Velocity $u$ (\code{float32})
    \item Vertical Velocity $v$ (\code{float32})
    \item Pressure $p$ (\code{float32})
    \item Cell Flags, defining which adjacent squares are boundaries (\code{uint8})
\end{enumerate}
This structure is somewhat unintuitive and error-prone, an example being the Cell Flags which may end up being inconsistent between adjacent cells, but it has been kept for the sake of compatibility with the original simulation.
